### 2025-03-08 causal-inference.qmd
20250220 - copyediting
20250227 - clean up
20250308 - R packages

However, it can be problematic to control for variables indiscriminantly [@Spector2010; @Wysocki2022].
The use of [causal diagrams](#sec-causalDiagrams) can inform which variables are important to be included as control variables, and—just as important—which variables not to include as control variables, as described in @sec-causalDiagrams.
Another approach that is sometimes used for matching, as a sample weight, or with stratification is the use of propensity scores.

@@

A key tool when describing a research question or hypothesis is to create a conceptual depiction of the hypothesized causal processes.
A causal diagram depicts the hypothesized causal processes that link two or more variables.
A common form of causal diagrams is the directed acyclic graph (DAG).
DAGs provide a helpful tool to communicate about causal questions and help identify how to avoid bias (i.e., overestimation) in associations between variables due to [confounding](#sec-causalDiagramConfounding) (i.e., common causes) [@Digitale2022].
For instance, from a DAG, it is possible to determine what variables it is important to control for (and how)—and just as importantly, what not to control for—in order to get unbiased estimates of the effect of one variable on another variable or of the association between two variables of interest.
To create DAGs, you can use the `R` package `dagitty` [@Textor2017] or the associated browser-based extension, DAGitty: <https://dagitty.net> (archived at <https://perma.cc/U9BY-VZE2>).
Examples of various causal diagrams that could explain why `X` is associated with `Y` are in Figures [-@fig-XCausesY], [-@fig-YCausesX] and [-@fig-ZCausesXandY].

@@

>>>>

However, it can be problematic to control for variables indiscriminantly [@Spector2010; @Wysocki2022].
The use of [causal diagrams](#sec-causalDiagrams) can inform which variables are important to be included as control variables, and—just as important—which variables not to include as control variables, as described in @sec-causalDiagrams.
Another approach that is sometimes used to control for variables—either by using matching, or as a sample weight, or with stratification—is the use of propensity scores.

@@

A key tool when describing a research question or hypothesis is to create a conceptual depiction of the hypothesized causal processes.
A causal diagram depicts the hypothesized causal processes that link two or more variables.
A common form of causal diagrams is the directed acyclic graph (DAG).
DAGs provide a helpful tool to communicate about causal questions and help identify how to avoid bias (i.e., overestimation) in associations between variables due to [confounding](#sec-causalDiagramConfounding) (i.e., common causes) [@Digitale2022].
For instance, from a DAG, it is possible to determine what variables it is important to control for (and how)—and just as importantly, what not to control for—in order to get unbiased estimates of the effect of one variable on another variable or of the association between two variables of interest.
To create DAGs, you can use the `R` package `dagitty` [@R-daggity; @Textor2016_packages] or the associated browser-based extension, DAGitty: <https://dagitty.net> (archived at <https://perma.cc/U9BY-VZE2>).
Examples of various causal diagrams that could explain why `X` is associated with `Y` are in Figures [-@fig-XCausesY], [-@fig-YCausesX] and [-@fig-ZCausesXandY].

@@

Here is an alternative way of specifying the same diagram (more similar to `lavaan` syntax) using the `ggdag` package [@R-ggdag]:

### 2025-03-08 cluster-analysis.qmd
20250308 - R packages

There are many approaches to cluster analysis, including model-based clustering, density-based clustering, centroid-based clustering, hierarchical clustering (aka connectivity-based clustering), etc.
An overview of approaches to cluster analysis in `R` is provided by @Kassambara2017a.
In this chapter, we focus on examples using model-based clustering with the `mclust` package [@Scrucca2023], which uses Gaussian finite mixture modeling.
The various types of `mclust` models are provided here:
<https://mclust-org.github.io/mclust/reference/mclustModelNames.html>.

>>>>

There are many approaches to cluster analysis, including model-based clustering, density-based clustering, centroid-based clustering, hierarchical clustering (aka connectivity-based clustering), etc.
An overview of approaches to cluster analysis in `R` is provided by @Kassambara2017a.
In this chapter, we focus on examples using model-based clustering with the `R` package `mclust` [@R-mclust; @Scrucca2023_packages], which uses Gaussian finite mixture modeling.
The various types of `mclust` models are provided here:
<https://mclust-org.github.io/mclust/reference/mclustModelNames.html>.

### 2025-03-08 data-visualization.qmd
20250308 - R packages

The Data Visualization Catalogue provides examples of various types of plots depending on one's goal: <https://datavizcatalogue.com/search.html>.
The R Graph Gallery provides examples of various types of plots and how to create them in `R`: <https://r-graph-gallery.com>.
Books on data visualization in `R` include [*`ggplot2`: Elegant Graphics for Data Analysis*](https://ggplot2-book.org) [@ggplot3rdEd] and [*`R` Graphics Cookbook: Practical Recipes for Visualizing Data*](https://r-graphics.org) [@Chang2018].
In this chapter, we will examine how to create statistical graphics to visualize data.
We will create the plots using the `ggplot2` package.
When creating plots in `ggplot2` with multiple points or lines (e.g., multiple players or levels of a predictor variable), it is easiest to do so with the data in [long form](#sec-wideToLong) (as opposed to [wide form](#sec-longToWide)).

@@

It is also important to use color schemes with distinguishable colors.
Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by ColorBrewer (<https://colorbrewer2.org>) and are available using the `scale_color_brewer()` and `scale_fill_brewer()` functions of the `ggplot2` package, as demonstrated in @fig-barPlotColorScheme (@sec-barPlotModifiedColorScheme).
There are a variety of resources for color schemes that are accessible to color-blind viewers:

- <https://www.datylon.com/blog/data-visualization-for-colorblind-readers> (archived at <https://perma.cc/7VTA-Y8YS>)
- <https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html> (archived at <https://perma.cc/NK9K-HL7L>)
- @Nunez2018

@@

A ridgeline plot of fantasy points is depicted in @fig-ridgeline1.

@@

Second, we create the base layer of the plot using the `ggplot()` function of the `ggplot2` package, as in @fig-scatterplot1.
We specify the data object and the variables in the data object that are associated with the x- and y-axes:

@@

Third, we create a scatterplot using the `geom_point()` function from the `ggplot2` package, as in @fig-scatterplot2:

@@

After creating our plot, we can make the plot interactive using the `ggplotly()` function from the `plotly` package, as in @fig-scatterplot11.

@@

After creating our plot, we can make the plot interactive using the `ggplotly()` function from the `plotly` package.

@@

We can also add a marginal density/histogram, as in @fig-scatterplot12:

@@

@@

Defensive EPA per play during the 2024 NFL season is depicted as a function of offensive EPA per play in @fig-plotDefOffEPA.

>>>>

The Data Visualization Catalogue provides examples of various types of plots depending on one's goal: <https://datavizcatalogue.com/search.html>.
The R Graph Gallery provides examples of various types of plots and how to create them in `R`: <https://r-graph-gallery.com>.
Books on data visualization in `R` include [*`ggplot2`: Elegant Graphics for Data Analysis*](https://ggplot2-book.org) [@ggplot3rdEd] and [*`R` Graphics Cookbook: Practical Recipes for Visualizing Data*](https://r-graphics.org) [@Chang2018].
In this chapter, we will examine how to create statistical graphics to visualize data.
We will create the plots using the `ggplot2` package [@R-ggplot2; @Wickham2016_packages].
When creating plots in `ggplot2` with multiple points or lines (e.g., multiple players or levels of a predictor variable), it is easiest to do so with the data in [long form](#sec-wideToLong) (as opposed to [wide form](#sec-longToWide)).

@@

It is also important to use color schemes with distinguishable colors.
Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by ColorBrewer (<https://colorbrewer2.org>) and are available using the `scale_color_brewer()` and `scale_fill_brewer()` functions of the `ggplot2` package [@R-ggplot2], as demonstrated in @fig-barPlotColorScheme (@sec-barPlotModifiedColorScheme).
There are a variety of resources for color schemes that are accessible to color-blind viewers:

- <https://www.datylon.com/blog/data-visualization-for-colorblind-readers> (archived at <https://perma.cc/7VTA-Y8YS>)
- the `viridis` package [@R-viridis; @Garnier2024a_packages]
  - <https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html> (archived at <https://perma.cc/NK9K-HL7L>)
- @Nunez2018

@@

A ridgeline plot of fantasy points is depicted in @fig-ridgeline1 using the `ggridges` package [@R-ggridges].

@@

Second, we create the base layer of the plot using the `ggplot()` function of the `ggplot2` package [@R-ggplot2], as in @fig-scatterplot1.
We specify the data object and the variables in the data object that are associated with the x- and y-axes:

@@

Third, we create a scatterplot using the `geom_point()` function from the `ggplot2` package [@R-ggplot2], as in @fig-scatterplot2:

@@

After creating our plot, we can make the plot interactive using the `ggplotly()` function from the `plotly` package [@R-plotly; @Sievert2020_packages], as in @fig-scatterplot11.

@@

After creating our plot, we can make the plot interactive using the `ggplotly()` function from the `plotly` package [@R-plotly; @Sievert2020_packages].

@@

We can also add a marginal density/histogram, as in @fig-scatterplot12 using the `ggExtra` package [@R-ggExtra]:

@@

We use the `gghighlight` package [@R-gghighlight] to highlight particular elements:

@@

Defensive EPA per play during the 2024 NFL season is depicted as a function of offensive EPA per play in @fig-plotDefOffEPA using the `nflplotR` package [@R-nflplotR].

### 2025-03-08 download-football-data.qmd
20250227 - clean up
20250308 - R packages

Below, we provide examples for how to download various types of NFL data using the `nflreadr` package.
For additional resources, @Congelio2023 provides a helpful introductory text for working with NFL data in `R`.
We save each data file after downloading it, so we can use the data in subsequent chapters.
If you have difficulty downloading the data files using the `nflreadr` package, we also saved the data files so they are publicly available on the Open Science Framework: <https://osf.io/z6pg4>.

@@

To download play-by-play data from prior weeks and seasons, we can use the `load_pbp()` function of the `nflreadr` package.
We add a progress bar using the `with_progress()` function from the `progressr` package because it takes a while to run.
A Data Dictionary for the play-by-play data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pbp.html>

@@

We can download historical week-by-week actual player statistics using the `load_player_stats()` function from the `nflreadr` package.
A Data Dictionary for statistics for offensive players is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats.html>.
A Data Dictionary for statistics for defensive players is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats_def.html>.

@@

First, we can compute the players' career statistics using the `calculate_stats()` function from the `nflfastR` package.

@@

First, we can compute the players' career statistics using the `calculate_player_stats()`, `calculate_player_stats_def()`, and `calculate_player_stats_kicking()` functions from the `nflfastR` package for offensive players, defensive players, and kickers, respectively.

@@

- Decide whether to use `calculate_stats()` from `nflfastR` or `load_player_stats()` from `nflreadr`.
- Save/update data file in repo with the data generated from the relevant code, and follow the data file through the rest of the book, updating code as necessary.

@@

>>>>

Below, we provide examples for how to download various types of NFL data using the `nflreadr` package [@R-nflreadr].
For additional resources, @Congelio2023 provides a helpful introductory text for working with NFL data in `R`.
We save each data file after downloading it, so we can use the data in subsequent chapters.
If you have difficulty downloading the data files using the `nflreadr` package [@R-nflreadr], we also saved the data files so they are publicly available on the Open Science Framework: <https://osf.io/z6pg4>.

@@

To download play-by-play data from prior weeks and seasons, we can use the `load_pbp()` function of the `nflreadr` package [@R-nflreadr].
We add a progress bar using the `with_progress()` function from the `progressr` package [@R-progressr] because it takes a while to run.
A Data Dictionary for the play-by-play data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pbp.html>

@@

We can download historical week-by-week actual player statistics using the `load_player_stats()` function from the `nflreadr` package [@R-nflreadr].
A Data Dictionary for statistics for offensive players is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats.html>.
A Data Dictionary for statistics for defensive players is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats_def.html>.

@@

First, we can compute the players' career statistics using the `calculate_stats()` function from the `nflfastR` package [@R-nflfastR].

@@

First, we can compute the players' career statistics using the `calculate_player_stats()`, `calculate_player_stats_def()`, and `calculate_player_stats_kicking()` functions from the `nflfastR` package [@R-nflfastR] for offensive players, defensive players, and kickers, respectively.

@@

- Decide whether to use `calculate_stats()` from `nflfastR` [@R-nflfastR] or `load_player_stats()` from `nflreadr` [@R-nflreadr].
- Save/update data file in repo with the data generated from the relevant code, and follow the data file through the rest of the book, updating code as necessary.

@@

We calculate the player's age based on the difference between dates using the `lubridate` package [@R-lubridate]:

### 2025-03-08 getting-started.qmd
20250226 - reprex do's and don'ts
20250227 - clean up
20250308 - R packages

The book uses `R` for statistical analyses (<http://www.r-project.org>).
R is a free software environment; you can download it at no charge here: <https://cran.r-project.org>.

@@

- Intro to `R`: <https://www.statmethods.net>
- Video training courses in `R` skills: <https://www.pluralsight.com/search?q=R>
- Browse the `Cookbook for R` to find solutions to common tasks and problems: <http://www.cookbook-r.com>
- Browse the `R Graph Gallery` to find examples of various graphs: <https://r-graph-gallery.com>
- Free `Codeacademy` course on `R`: <https://www.codecademy.com/learn/learn-r>
- Free `Coursera` courses on `R`: <https://www.coursera.org/search?query=R>
- Watch these videos from `Coursera`: <https://blog.revolutionanalytics.com/2012/12/coursera-videos.html>
- `Posit`/`Rstudio` Webinars: <https://posit.co/resources/videos/>
- UCLA Stats Website: <https://stats.idre.ucla.edu/r/>
- Introduction to `R` course on `Datacamp`: <https://www.datacamp.com/courses/free-introduction-to-r>
- Teaching `R` in a Kinder, Gentler, More Effective Manner: <https://github.com/matloff/TidyverseSkeptic>
- Learn `R` interactively with `swirl`: <https://swirlstats.com>
- Use the `learnr` package: <https://rstudio.github.io/learnr/>
- Resources for learning `tidyverse`, which is a collection of `R` packages for data management: <https://www.tidyverse.org/learn/>
- You will sometimes find relevant articles on `R-bloggers`: <https://www.r-bloggers.com>

@@

- Read the posting guidelines before posting!
- Be respectful of other people and their time.
`R` is free software.
People are offering their free time to help.
They are under no obligation to help you.
If you are disrespectful or act like they owe you anything, you will rub people the wrong way and will be less likely to get help.
- Provide a minimal, reproducible example.
Providing a minimal, reproducible example can be crucial for getting a helpful response.
By going to the trouble of creating a minimal, reproducible example and identifying the minimum conditions necessary to reproduce the issue, you will often figure out how to resolve it.
Here are guidelines on providing a minimal, reproducible example: <https://stackoverflow.com/help/minimal-reproducible-example> (archived at <https://perma.cc/6NUB-UTYF>).
Here are a good example and guidelines for providing a minimal, reproducible example in `R`: <https://stackoverflow.com/a/5963610> (archived at <https://perma.cc/PC9L-DQZG>).
My strong recommendation is to provide a `reprex` whenever possible: <https://reprex.tidyverse.org>.

@@

   install.packages(
     c("petersenlab","remotes","nflreadr","nflfastR","nfl4th","nflplotR",
     "gsisdecoder","progressr","lubridate","tidyverse","psych"))
   ```
1. Some necessary packages, including the `ffanalytics` package, are hosted in GitHub and need to be installed using the following code (after installing the `remotes` package above):
   ```{r}
   #| eval: false

@@

For instance, you can use the following code to install the `tidyverse` package:

>>>>

The book uses the software `R` [@R-base] for statistical analyses (<http://www.r-project.org>).
R is a free software environment; you can download it at no charge here: <https://cran.r-project.org>.

@@

- Intro to `R`: <https://www.statmethods.net>
- Video training courses in `R` skills: <https://www.pluralsight.com/search?q=R>
- Browse the `Cookbook for R` to find solutions to common tasks and problems: <http://www.cookbook-r.com>
- Browse the `R Graph Gallery` to find examples of various graphs: <https://r-graph-gallery.com>
- Free `Codeacademy` course on `R`: <https://www.codecademy.com/learn/learn-r>
- Free `Coursera` courses on `R`: <https://www.coursera.org/search?query=R>
- Watch these videos from `Coursera`: <https://blog.revolutionanalytics.com/2012/12/coursera-videos.html>
- `Posit`/`Rstudio` Webinars: <https://posit.co/resources/videos/>
- UCLA Stats Website: <https://stats.idre.ucla.edu/r/>
- Introduction to `R` course on `Datacamp`: <https://www.datacamp.com/courses/free-introduction-to-r>
- Teaching `R` in a Kinder, Gentler, More Effective Manner: <https://github.com/matloff/TidyverseSkeptic>
- Learn `R` interactively with `swirl`: <https://swirlstats.com>
- Use the `learnr` package [@R-learnr]: <https://rstudio.github.io/learnr/>
- Resources for learning `tidyverse` [@R-tidyverse; @Wickham2019_packages], which is a collection of `R` packages for data management: <https://www.tidyverse.org/learn/>
- You will sometimes find relevant articles on `R-bloggers`: <https://www.r-bloggers.com>

@@

- Read the posting guidelines before posting!
- Be respectful of other people and their time.
`R` is free software.
People are offering their free time to help.
They are under no obligation to help you.
If you are disrespectful or act like they owe you anything, you will rub people the wrong way and will be less likely to get help.
- Provide a minimal, reproducible example.
Providing a minimal, reproducible example can be crucial for getting a helpful response.
By going to the trouble of creating a minimal, reproducible example and identifying the minimum conditions necessary to reproduce the issue, you will often figure out how to resolve it.
Here are guidelines on providing a minimal, reproducible example: <https://stackoverflow.com/help/minimal-reproducible-example> (archived at <https://perma.cc/6NUB-UTYF>).
Here are a good example and guidelines for providing a minimal, reproducible example in `R`: <https://stackoverflow.com/a/5963610> (archived at <https://perma.cc/PC9L-DQZG>).
My strong recommendation is to provide a `reprex` whenever possible: <https://reprex.tidyverse.org>.
Here are Reprex do's and don'ts: <https://reprex.tidyverse.org/articles/reprex-dos-and-donts.html>.

@@

   install.packages(
     c("petersenlab","remotes","nflreadr","nflfastR","nfl4th","nflplotR",
     "gsisdecoder","progressr","lubridate","tidyverse","psych"))
   ```
1. Some necessary packages, including the [`ffanalytics` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], are hosted in GitHub and need to be installed using the following code (after installing the `remotes` package [@R-remotes] above):
   ```{r}
   #| eval: false

@@

For instance, you can use the following code to install the `tidyverse` package [@R-tidyverse]:

### 2025-03-08 index.qmd
20250308 - R packages

# Automatically create a bib database for R packages
packagesForBibliography <- c(
  "petersenlab", "ffanalytics", "pROC", "rms"
)

>>>>

# Automatically create a bib database for R packages
packagesForBibliography <- c(
  "petersenlab", "ffanalytics",
  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "downlit", "xml2",
  "nflreadr", "nflplotR", "nflfastR", "gsisdecoder", "progressr", "DescTools",
  "pwr", "pwrss", "WebPower", "XICOR", "dagitty", "ggdag", "lubridate",
  "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", "pROC", "lme4",
  "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", "AICcmodavg", "rstan",
  "brms", "bbmle", "fitdistrplus", "mclust", "magrittr", "viridis", "plotly",
  "quantmod", "fPortfolio", "xts", "zoo", "forecast", "stringi", #"webshot2", 
  "parallel", "parallelly", "ggridges", "cmdstanr",
  "learnr", "remotes" #"mgcv",

)

### 2025-03-08 intro.qmd
20240915 - videos
20241016 - render the figure caption as alt text
20250308 - R packages

The book provides data analysis examples using the statistical analysis software, `R`.
Why `R`?

- `R` is free!
Anyone can use it.
- `R` is open source—it is not a black box.
You can see what is going on "under the hood" and can examine the code for any function or computation you perform.
You can even modify and improve these functions by changing the code, and you can create your own functions.
- `R` is open platform—you can use it on multiple platforms, including Windows, MacOS, and Linux.
- `R` has advanced statistics capabilities.
It was designed for statistical analysis and has strong capabalities for data wrangling.
- `R` has capabilities for state-of-the-art graphics.
It has advanced capabilities for creating statistical graphics.
- `R` is widely used—there is a large community of people who use `R` for data analysis that you can draw upon for help from others.
- `R` analyses are based on code (rather than a graphical user interface), which allows reproducibility—with the same data, code, and setup (platform, `R` version, package versions, etc.), you should get the same answer every time.
There are strong resources available for ensuring your analyses in `R` are reproducible by others [@Gandrud2020].
- Anyone (including you) can contribute `R` packages to the community to improve its functionality.
Statistical experts from all over the world have contributed open source packages to `R` for specialized tasks.
In the chance there is not an `R` package that does what you need to do, you can write a function to perform the task and can contribute it as a package to the community for others to use and improve.
The number of `R` packages contributed to the community is growing at a rapid rate.
As of this writing, over 20,000 packages have been contributed to the [Comprehensive R Archive Network](https://cran.r-project.org/web/packages/) (CRAN).
And many more are stored on publicly available version control repositories like GitHub and GitLab.
Chances are, if there is an analysis you need to do, an `R` package exists to do it.

@@

![Video of Prof. Benjamin Motz Discussing Statistics and His Fantasy Football Course.](images/motzVideo.mp4){width=100%}

>>>>

The book provides data analysis examples using the statistical analysis software, `R` [@R-base].
Why `R`?

- `R` is free!
Anyone can use it.
- `R` is open source—it is not a black box.
You can see what is going on "under the hood" and can examine the code for any function or computation you perform.
You can even modify and improve these functions by changing the code, and you can create your own functions.
- `R` is open platform—you can use it on multiple platforms, including Windows, MacOS, and Linux.
- `R` has advanced statistics capabilities.
It was designed for statistical analysis and has strong capabilities for data wrangling.
- `R` has capabilities for state-of-the-art graphics.
It has advanced capabilities for creating statistical graphics.
- `R` is widely used—there is a large community of people who use `R` for data analysis that you can draw upon for help from others.
- `R` analyses are based on code (rather than a graphical user interface), which allows reproducibility—with the same data, code, and setup (platform, `R` version, package versions, etc.), you should get the same answer every time.
There are strong resources available for ensuring your analyses in `R` are reproducible by others [@Gandrud2020].
- Anyone (including you) can contribute `R` packages to the community to improve its functionality.
Statistical experts from all over the world have contributed open source packages to `R` for specialized tasks.
In the chance there is not an `R` package that does what you need to do, you can write a function to perform the task and can contribute it as a package to the community for others to use and improve.
The number of `R` packages contributed to the community is growing at a rapid rate.
As of this writing, over 20,000 packages have been contributed to the [Comprehensive R Archive Network](https://cran.r-project.org/web/packages/) (CRAN).
And many more are stored on publicly available version control repositories like GitHub and GitLab.
Chances are, if there is an analysis you need to do, an `R` package exists to do it.

@@

![Video of Prof. Benjamin Motz Discussing Statistics and His Fantasy Football Course. From: <https://www.facebook.com/watch/?v=10155572257183615>.](images/motzVideo.mp4){width=100%}\

### 2025-03-08 mixed-models.qmd
20250308 - R packages

@@

@@

@@

@@

@@

@@

Information about smooth terms in the `mgcv` package is provided at the following link: <https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/smooth.terms.html>.

@@

>>>>

We can estimate model fit using the `MuMIn` package [@R-MuMIn]:

@@

We compare models using the `bbmle` [@R-bbmle] and `MuMIn` [@R-MuMIn] packages.

@@

We estimate the multilevel models using the `lmerTest` package [@R-lmerTest], which is an extension of the `lme4` package [@R-lme4; @Bates2015_packages]:

@@

We use the `emmeans` package [@R-emmeans] to determine the model-implied number of points per season by position.

@@

We compute model fit using the using the `AICcmodavg` [@R-AICcmodavg], `bbmle` [@R-bbmle], and `MuMIn` [@R-MuMIn] packages

@@

We can use the `fitdistrplus` package [@R-fitdistrplus; @DelignetteMuller2015_packages]:

@@

Information about smooth terms in the `mgcv` package [@R-mgcv; @Wood2017] is provided at the following link: <https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/smooth.terms.html>.
We estimate the model using the `brms` package [@R-brms; @Buerkner2017_packages; @Buerkner2018_packages], which provides an interface to the `rstan` package [@R-rstan].

@@

We use the `cmdstanr` backend [@R-cmdstanr] and threading using the `parellely` package [@R-parallelly] for parallel (faster) processing:
